Relationship between solar wind, $D_{st}$, and plasmasphere trough mass density on one-hour time scales

### 2016-02-26 paper.tex
Introduction additions

\section{Introduction}
\cite{Gallagher1988EmpiricalModelPlasmasphere} \vinote{Too far back?} first models the the plasmasphere as a simple linear fit dependent on L shell, geographic height, geomagnetic latitude, and $K_p$. \cite{Hamilton1988RingCurrentDevelopment} simultaneously analyzes a strong geomagnetic storm by comparing empirical energy measurements of the ring current to $D_{st}$, along with measuring ring current compositions before, during, and after the storm.

\cite{Lotoaniu1999PlasmaMassDensity} uses CRRES to compare in-situ measured plasma mass density to that inferred by ground-based ULF wave measurements extended to the plasmatrough.

@@

\cite{Takahashi2010} developed a mass density dataset using measurements from the GOES satellites over the years 1980 through 1992, with most measurements in the range $L=6.8\pm0.2 R_E$. The mass density $\rho$ was estimated using the the Alfven wave velocity relationship, $V_A=B/\sqrt{\mu_0\rho}$, a magnetic field model, and a numerical solution to a wave equation with an ionospheric boundary and the assumption of a zero resistance ionosphere.  The equatorial mass density, $\rho_{eq}$, was derived from the estimated mass density using a power law dependence on the geocentric distance to the field line of the observation, $R$, $\rho=\rho_{eq}L/(R/R_E)$.

\cite{Takahashi2010} \vinote{Necessary repetition?} noted that downward spikes in the $D_{st}$ index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8~$R_E$. For five storms over a 20-day period, two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, an enhancment in $\rho_{eq}$ appeared on the same day as minimum $D_{st}$, indicating the possibility of a solar wind control over $\rho_{eq}$.

@@

%\cite{Denton2006} show how $D_{st}$ affects the distribution of plasma density along different magnetic latitudes, and specifically along the same field lines as looked at in later papers (6-8$R_E$). Though this shows that the trends for density may differ between field lines, it's mentioned mostly as a point for future research as the data used in this paper is already adjusted for one field line, as described by \cite{Takahashi2010}.

%\cite{Lavraud2006} show that storms preceded by a northward interplanetary magnetic field (IMF) tend to have higher plasma density than those preceded by a southward IMF, and suggest an increased geoeffectiveness from a northward IMF and dense plasma sheet.

%\cite{Tsurutani1997} show how the density of a corotating interaction region (CIR) shows strong preconditioning for a drop in $D_{st}$ and spike in magnetic field intensity. Assuming a reasonable amount of coupling between that region's density and the plasmaspheric density, this result could be relatable to others focusing on the plasmasphere.

>>>>

\section{Introduction}

The plasma trough is the transition region between the high density and low temperature plasmasphere and the lower density and, in the equatorial plane, higher energy magnetosphere \citep{Mayr1987}.  Although there are well--validated models of the plasmasphere density as a function of $L$ (\cite{Gallagher1988EmpiricalModelPlasmasphere}, \cite{Lemaire}, \cite{Takahashi2006}, and \cite{Takahashi2010}) and plasmapause boundaries (\cite{Lemaire}, \cite{Moldwin}, \cite{OBrien}), the dependence of the plasma trough on solar wind conditions has not been extensively studied.

\cite{Gallagher1988EmpiricalModelPlasmasphere} modeled the the plasmasphere as a simple linear fit dependent on $L$ shell, geographic altitude, geomagnetic latitude, and $K_p$.  \vnote{What was major finding?} \cite{Hamilton1988RingCurrentDevelopment} analyzed a strong geomagnetic storm and compared energy measurements of the ring current to $D_{st}$, along with measurements of ring current compositions before, during, and after the storm.  \vnote{What did they find/conclude?}

\cite{Lotoaniu1999PlasmaMassDensity} used measurements from the \vnote{???} instrument on the CRRES sateelite and compared in-situ measured plasma mass density to that inferred by ground-based ULF wave measurements that extended to the plasmatrough.  \vnote{What did they find/conclude?}

@@

\cite{Denton2006} showed a relationship between $D_{st}$ and the field line distribution of plasma density, and specifically along the same field lines as looked at in later papers (6-8$R_E$). They found that the distribution depends on $L$,.

\cite{Takahashi2010} developed a mass density dataset using measurements from the SEM \vnote{Check instrument} instruments on the GOES satellites from 1980 through 1992, with most measurements in the range $L=6.8\pm0.2 R_E$. The mass density $\rho$ was estimated using the the Alf\v'en wave velocity relationship, $V_A=B/\sqrt{\mu_0\rho}$, a magnetic field model, and a numerical solution to a wave equation with an ionospheric boundary and the assumption of a zero resistance ionosphere.  The equatorial mass density, $\rho_{eq}$, was derived from the estimated mass density using a power law dependence on the geocentric distance to the field line of the observation, $R$, $\rho=\rho_{eq}L/(R/R_E)$.

\cite{Takahashi2010} noted that downward spikes in the $D_{st}$ index coincided with with significant changes in $\rho_{eq}$ at for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, an enhancment in $\rho_{eq}$ appeared on the same day as minimum $D_{st}$, indicating the possibility of a solar wind control over $\rho_{eq}$.

@@

In this work we consider the dependence of plasmaspheric mass density in the \cite{Takahashi2010} in more detail in an attempt to identify the solar wind processes that drives changes in the plasma trough density.
